Spring haul-out behavior of seals in the Bering and Chukchi seas

### 2024-05-29 analysis/paper/London_HauloutBehavior.Rmd
move permit details to methods section

For this study we used haul-out behavior data and location estimates from
bio-loggers deployed on bearded, ribbon, and spotted seals in the Bering,
Chukchi, and western Beaufort seas by multiple organizations as part of
collaborative investigations from 2005 through 2021. Seals were captured using
nets and bio-loggers were attached during studies based in coastal communities
or on research ships. Ship-based capture events occurred during spring near the
southern ice edge in the Bering Sea between 2005 and 2018. Land-based capture
events occurred between 2005 and 2020 from May to October, generally between the
Alaska coastal communities of Scammon Bay in the Bering Sea, Utqiaġvik
in the Chukchi Sea, and Nuiqsut in the Beaufort Sea (Supplemental
Material, [S1](#s1)). Data from additional
deployments along the Kamchatka peninsula in the western Bering Sea are also
included. We refer readers to Figure \@ref(fig:deployplot) and the primary literature
for detailed capture and bio-logger attachment methods (see also Supplemental
Material, [S1](#s1)).

@@

```{r deployplot, include=TRUE, fig.asp = 0.4, fig.cap = "(ref:deployMapCaption)"}
targets::tar_load(deploy_details, store = here::here("_targets"))
targets::tar_load(survey_area, store = here::here("_targets"))

@@

p <- ggplot() +
  annotation_spatial(survey_area, color = "#A26E7C", 
                fill = "#A26E7C", 
                alpha = 0.15,
                linewidth = 0.35) +
  layer_spatial(deploy_sf, aes(fill=species),
                linewidth=0) +
  annotation_spatial(world,
                     linewidth=0,
                     fill = "grey80") +
  geom_sf_text(data = label_coords_bering, aes(label = "Bering sea"), size = 2, color = "grey30") +
  geom_sf_text(data = label_coords_chukchi, 
               aes(label = "Chuckchi \nsea"), size = 2, lineheight = .85, color = "grey30") +
  geom_sf_text(data = label_coords_beaufort, 
               aes(label = "Beaufort \nsea"), size = 2, lineheight = .85, color = "grey30") +
  facet_grid(cols = vars(species), labeller = species_labels) +
  scale_fill_manual(values = species_colors,
                    aesthetics = "fill", 
                    name = "species") + 
  labs(caption = "<span style = 'color: #A26E7C;'><span style = 'font-family: \"Font Awesome 6 Free Regular\";'>&#xf0c8;</span> spatial extent of aerial survey effort in US and Russian waters</span>") +
  theme(legend.position = "none",
        axis.title = element_blank(),
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.x = element_markdown(face = "bold")
  )

@@

ggplot() +
  annotation_spatial(world,
                     linewidth = 0,
                     fill = "grey80") +
  layer_spatial(
    data = locs_count,
    aes(fill = ct),
    linewidth = 0.1,
    color = "grey75"
  ) +
  scico::scale_fill_scico(
    palette = "imola",
    trans = "log10",
    aesthetics = "fill",
    name = "number of haul-out records",
    guide = guide_colorbar(
      title.position = "bottom",
      barwidth = 15,
      barheight = 0.5,
      title.hjust = 0.5
    )
  ) +
  facet_grid( ~ species) +
  labs(
    caption = "One hexagonal cell represents 50 km^2^"
  ) +
  theme(
    legend.position = "bottom"
  )
```

@@

The findings and conclusions in the paper are those of the author(s) and do not
necessarily represent the views of the National Marine Fisheries Service, NOAA.
Any use of trade, product, or firm names does not imply an endorsement by the
U.S. Government. Funding for this study was provided by the U.S. National
Oceanic and Atmospheric Administration. The field work was conducted under the
authority of Marine Mammal Protection Act Research Permits Nos. 782-1676,
782-1765, 15126, and 19309 issued by the National Marine Fisheries Service, and
Letters of Assurance of Compliance with Animal Welfare Act regulations, Nos.
A/NW 2010-3 and A/NW 2016-1 from the Alaska Fisheries Science Center/Northwest
Fisheries Science Center Institutional Animal Care and Use Committee (IACUC). ).
Funding to ADF&G for tagging seals was provided by the Bureau of Ocean Energy
Management (No. M13PC0015 for work in Kotzebue, Alaska in 2009) and the Office
of Naval Research (No. N00014-16-1-3019). ADF&G and NSB field work was covered by
Research Permits Nos. 358-1585, 358-1787, 15324, and 20466 and by ADF&G IACUC
permits Nos. 06-16, 09-21, 2014-03, 2015-25, 2016-23, 0027-2017-27,
0027-2018-29, 0027-2019-041.

>>>>

For this study we used haul-out behavior data and location estimates from
bio-loggers deployed on bearded, ribbon, and spotted seals in the Bering,
Chukchi, and western Beaufort seas by multiple organizations as part of
collaborative investigations from 2005 through 2021. Seals were captured using
nets and bio-loggers were attached during studies based in coastal communities
or on research ships. Ship-based capture events occurred during spring near the
southern ice edge in the Bering Sea between 2005 and 2018. Land-based capture
events occurred between 2005 and 2020 from May to October, generally between the
Alaska coastal communities of Scammon Bay in the Bering Sea, Utqiaġvik in the
Chukchi Sea, and Nuiqsut in the Beaufort Sea (Supplemental Material, [S1](#s1)).
Data from additional deployments along the Kamchatka peninsula in the western
Bering Sea are also included. We refer readers to Figure \@ref(fig:deployplot)
and the primary literature for detailed capture and bio-logger attachment
methods (see also Supplemental Material, [S1](#s1)). NOAA-led research was
conducted under the authority of Marine Mammal Protection Act Research Permits
Nos. 782-1676, 782-1765, 15126, and 19309 issued by the National Marine
Fisheries Service, and Letters of Assurance of Compliance with Animal Welfare
Act regulations, Nos. A/NW 2010-3 and A/NW 2016-1 from the Alaska Fisheries
Science Center/Northwest Fisheries Science Center Institutional Animal Care and
Use Committee (IACUC). ADF&G and NSB field work was covered by Research Permits
Nos. 358-1585, 358-1787, 15324, and 20466 and by ADF&G IACUC permits Nos. 06-16,
09-21, 2014-03, 2015-25, 2016-23, 0027-2017-27, 0027-2018-29, 0027-2019-041.

@@

```{r deployplot, include=TRUE, fig.asp = 0.42, fig.cap = "(ref:deployMapCaption)"}
targets::tar_load(deploy_details, store = here::here("_targets"))
targets::tar_load(survey_area, store = here::here("_targets"))

@@

p <- ggplot() +
  annotation_spatial(survey_area, color = "#A26E7C", 
                fill = "#A26E7C", 
                alpha = 0.15,
                linewidth = 0.35) +
  layer_spatial(deploy_sf, aes(fill=species),
                linewidth=0) +
  annotation_spatial(world,
                     linewidth=0,
                     fill = "grey80") +
  geom_sf_text(data = label_coords_bering, aes(label = "Bering sea"), size = 2, color = "grey30") +
  geom_sf_text(data = label_coords_chukchi, 
               aes(label = "Chuckchi \nsea"), size = 2, lineheight = .85, color = "grey30") +
  geom_sf_text(data = label_coords_beaufort, 
               aes(label = "Beaufort \nsea"), size = 2, lineheight = .85, color = "grey30") +
  facet_grid(cols = vars(species), labeller = species_labels) +
  scale_fill_manual(values = species_colors,
                    aesthetics = "fill", 
                    name = "species") + 
  labs(caption = "<span style = 'color: #A26E7C;'><span style = 'font-family: \"Font Awesome 6 Free Regular\";'>&#xf0c8;</span> spatial extent of aerial survey effort in US and Russian waters</span><br />world land vector data from naturalearthdata.com") +
  theme(legend.position = "none",
        axis.title = element_blank(),
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.x = element_markdown(face = "bold")
  )

@@

ggplot() +
  annotation_spatial(world,
                     linewidth = 0,
                     fill = "grey80") +
  layer_spatial(
    data = locs_count,
    aes(fill = ct),
    linewidth = 0.1,
    color = "grey75"
  ) +
  scico::scale_fill_scico(
    palette = "imola",
    trans = "log10",
    aesthetics = "fill",
    name = "number of haul-out records",
    guide = guide_colorbar(
      title.position = "bottom",
      barwidth = 15,
      barheight = 0.5,
      title.hjust = 0.5
    )
  ) +
  facet_grid( ~ species) +
  labs(
    caption = "One hexagonal cell represents 50 km^2^<br />world land vector data from naturalearthdata.com"
  ) +
  theme(
    legend.position = "bottom"
  )
```

@@

The findings and conclusions in the paper are those of the author(s) and do not
necessarily represent the views of the National Marine Fisheries Service, NOAA.
Any use of trade, product, or firm names does not imply an endorsement by the
U.S. Government. Funding for this study was provided by the U.S. National
Oceanic and Atmospheric Administration. Funding to ADF&G for tagging seals was provided by the Bureau of Ocean Energy
Management (No. M13PC0015 for work in Kotzebue, Alaska in 2009) and the Office
of Naval Research (No. N00014-16-1-3019).
